Reference model in feature selection problems \vspace{.1in}

### 2019-05-23 paper/ref_approach.tex
expanded introduction with references to reference model approach + motivation of interest in the paper

In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection. From a full Bayesian perspective, when the goal of the analysis is to do prediction, no selection should be carried out and integration over all parameters uncertainties, i.e. posterior distributions, is the recommended procedure. However, a high dimensional parameter space brings different difficulties as computation burden and overfitting. For such reason different feature selection procedures have been proposed both in the frequentist and Bayesian setting (e.g. see \cite{paper:feature_selection}, \cite{paper:vehtari_ojanen} and \cite{paper:model_selection}). Sometimes the task of the analysis is not prediction, but rather inference on which variables are significantly related to the phenomenon of interest. A common case is microarray data in medicine or biology, where for example for a very large number of genes, expression levels are measured for two sets of patients labeled as \textit{positive} or \textit{negative} to some disease (typically a \textit{small n large p} scenario).  The main goal of the analysis is to spot which subset of these genes is significantly related to the disease in order to perform further experiments. A possible way to proceed is to perform hypothesis testing to assess the significance of each gene, based on some modelling assumptions. \cite{book:efron} refers to this as multiple hypothesis testing problem. Such example could be still considered as feature selection, intended this time as the problem of ``identifying \textit{all} features that are predictive about (that is, statistically related to) the target variable'' \citep{paper:projpred}.

Regardless which kind of selection problem we are trying to solve, we consider examples of the reference model approach all those methods that make use of a full-encompassing model, i.e. the reference one, to carry out the selection. The idea of using a reference model is already present in some works from the 60s as \cite{paper:reference_lindley}. It is also the basis for the projection predictive approach \citep{paper:original_proj, paper:projpred}, from which we have taken the hint of using the reference model as a filter (see Section \ref{reference-model-approach}).

\begin{itemize}
\item add original contribution that are:
\item motivation and example on how the use of a reference model improves the selection (projpred example)
\item a comparison between using a reference model or not in the normal means problem  both with simulated data and the bodyfat data
\end{itemize}

@@

\hypertarget{reference-model-approach}{%
\section{Reference model approach to feature selection}\label{reference-model-approach}}

The reference model approach has his roots in  the $\mathcal{M}$-\textit{complete} framework \citep{book:bernardo_smith,paper:vehtari_ojanen} and started to be used in the feature selection context already starting from the 60s with \cite{paper:reference_lindley}. There is not a specific definition of what a reference model is; ideally any full-encompassing model which has a good predictive performance is a candidate reference model. General guidelines on how to devise it are the same about any model in a data analysis context, expect for the fact that no selection is carried out for the reference model. \cite{paper:projpred} showed that through the projection predictive approach the reference model can helps in terms of stability and model size in the selection process. In this Section we aim to give some hints on why this happens and on how we can generally rely on a reference model approach to feature selection with a simple example.

@@

What is shown in Figure \ref{fig:correlation} is related also to the projection predictive approach. In case of a submodel belonging to the generalised linear model family, it happens that the draw-by-draw projection coincide with the maximum likelihood estimator with observed data substituted with reference model's predictions \citep{paper:projpred}. It is possible to see the idea of the reference model as a filter on our data also from the projection of the predictive distribution itself. Let us call $p_{\text{ref}}(\tilde{y}|D)$ the posterior predictive distribution for a new observation $\tilde{y}$ for the reference model and let $\pi(\theta)$ and $q_{\pi}(\tilde{y})$ be respectively the submodel parameter and the induced predictive distributions. The projection of the predictive distribution allows to find the optimal parameter distribution $\pi^{*}$ minimising the KL discrepancy among the induced predictive distribution of the submodel and the reference model one:
\
\begin{equation} 
\pi^{*}\quad \text{s.t.} \quad \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi^{*}}(\tilde{y})]=\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
Consider now the predictive performance of such submodel, a common utility in Bayesian context is the expected log-predictive density defined as:
\
\begin{equation} 
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
\end{equation}
where $p_{t}$ is the distribution of a new observation under the true data generation mechanism. Maximising the elpd of a model is equivalent to minimising the KL-divergence of its predictive distribution from the true data generation mechanism. Therefore the projection corresponds to maximising the elpd using as distribution of the true data generation mechanism the posterior predictive distribution of the reference model, i.e.:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] \leftrightarrow \underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p_{\text{ref}}(\tilde{y}|D)d\tilde{y} 
\end{equation}
Equation \eqref{proj_as_filter} clearly shows the action of the reference model as a filter on our target observations, substituting the unknown true data generation mechanism, that is usually estimated with the data empirical distribution, with the best known approximation that is available, i.e. the reference model posterior predictive distribution.

>>>>

In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection, that can be seen as a special case of model reduction. From a full Bayesian perspective, when the goal of the analysis is to do prediction, no selection should be carried out and integration over all parameters uncertainties, i.e. posterior distributions, is the recommended procedure. However, a high dimensional parameter space brings different difficulties as computation burden and overfitting. For such reason different feature selection procedures have been proposed both in the frequentist and Bayesian setting (e.g. see \cite{paper:feature_selection}, \cite{paper:vehtari_ojanen} and \cite{paper:model_selection}). Sometimes the task of the analysis is not prediction, but rather inference on which variables are significantly related to the phenomenon of interest. A common case is microarray data in medicine or biology, where for example for a very large number of genes, expression levels are measured for two sets of patients labeled as \textit{positive} or \textit{negative} to some disease (typically a \textit{small n large p} scenario).  The main goal of the analysis is to spot which subset of these genes is significantly related to the disease in order to perform further experiments. A possible way to proceed is to perform hypothesis testing to assess the significance of each gene, based on some modelling assumptions. \cite{book:efron} refers to this as multiple hypothesis testing problem. Such example could be still considered as feature selection, intended this time as the problem of ``identifying \textit{all} features that are predictive about (that is, statistically related to) the target variable'' \citep{paper:projpred}.

Regardless which kind of selection problem we are trying to solve, we consider examples of the reference model approach all those methods that make use of a full-encompassing model, i.e. the reference one. Such an idea is not new to the model reduction literature, it has been used for many years using different type of names. \cite{book:harrell} refers to the reference model as a full model that can be thought as the ``gold standard'' and shows some examples of how it can be used to look for a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximate the best model and the inheritance of the shrinkage when it is applied to the full model. \cite{paper:faraggi_nn} deal with the necessity of identifying interpretable risk groups in the context of survival data. Neural networks typically perform very well in terms of prediction, but it is not very clear how to interpret the significance of each covariate. \cite{paper:faraggi_nn} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups. \cite{paper:paul_preconditioning}, using the term preconditioning, explore approximating models fitting Lasso \citep{paper:tibshirani_lasso} or stepwise regression against consistent estimates $\hat{y}$ of a reference model instead of the observed response quantities $y$. \cite{book:harrell} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.
\\
The reference model approach has been used for long time also in the Bayesian framework, an example the pioneering work by \cite{paper:reference_lindley}. In the case of generalised linear models (GLMs), \cite{paper:tran_predictivelasso} provide a procedure to estimate and do variable selection with the goal of prediction. The best approximating model is recovered by optimising the Kullback-Liebler divergence between the predictive distribution of the reference model and the approximating model plus an $l_{1}$ penalty on the submodel coefficients. The resulting problem is convex and thus many efficient optimisation algorithm are available. Following \cite{paper:goutis_projection}, \cite{paper:original_proj}  instead project the posterior distribution of the reference model in each candidate model and then select the smallest one with enough explanatory power avoiding any prior elicitation for the submodels. In the special case of candidate models in the family of the GLMs, the optimisation problem correspond to the maximum likelihood estimation problem having the expected model point prediction by the reference model instead of the observed target values. Different extensions have followed \cite{paper:original_proj}, as \cite{paper:nott_projection} that add an $l_{1}$ penalty to select the best approximating model. \cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance with the reference model one.

\cite{paper:projpred} show in their experiments the benefits of the projection predictive approach in terms of predictive ability of the selected model and smaller size compared to other methods. It could be of interest to know how often a certain method includes non-relevant features. One could argue that such question can not have a general answer, because it depends both on the data and on the choice of the reference model. Moreover in real settings it is hard to define what relevant means, in most of the cases all the features have some level of relevance in term of prediction ability and thus the question looses of meaning.  However, it is still possible to simulate certain artificial scenarios and control which features are included. \cite{paper:model_selection} report results about the proportion of relevant and non-relevant features selected with different variable selection methods, showing overall better results in favour of the projection predictive approach. The explored submodels using the projection show also a higher performance and stability with respect to the compared methods, given the same submodel size.

In this paper we try to give further examples and intuitions on why the reference model helps. Using a real world dataset, we show the improvements in the selection when using the projection predictive approach versus the stepwise regression. Wherever it has been used, the reference model can be always seen acting as a filter on the observed data. Following this idea, we devise a simple reference model approach that can be applied on top of any selection procedure. Our goal is to show the reference model benefits, regardless of what specific procedure is applied. We take as reference problem the normal means estimation problem. We apply different state-of-the-art methods comparing in each case the results using the data or using a reference model. In case of full Bayesian methods, we report results in terms of shrinkage and accuracy of the estimates, whereas when a full selection is carried out, we compute quantities as false discovery rate, sensitivity and a measure of the stability of the selection. We believe that these results should motivate the use of a reference model when doing model reduction, regardless of the specific method used and thus we hope this work can be of interest to a large number of people.

@@

\hypertarget{reference-model-approach}{%
\section{Why the reference model helps}\label{reference-model-approach}}

The main assumption under any reference model approach is to operate in  the $\mathcal{M}$-\textit{complete} framework \citep{book:bernardo_smith,paper:vehtari_ojanen}, thus to be able to have a model, that is the reference model, which is the best available description of a future observation. There is not a specific definition of how a reference model should be; in practice, any full-encompassing model which has a good predictive performance is a candidate reference model. General guidelines on how to devise it are the same about any model in a data analysis context, expect for the fact that no selection is carried out for the reference model.

@@

What is shown in Figure \ref{fig:correlation} is related also to the projection predictive approach. In case of a submodel belonging to the generalised linear model family, it happens that the draw-by-draw projection coincide with the maximum likelihood estimator with observed data substituted with reference model's predictions \citep{paper:original_proj}. It is possible to see the idea of the reference model as a filter on our data also from the projection of the predictive distribution itself. Let us call $p_{\text{ref}}(\tilde{y}|D)$ the posterior predictive distribution for a new observation $\tilde{y}$ for the reference model and let $\pi(\theta)$ and $q_{\pi}(\tilde{y})$ be respectively the submodel parameter and the induced predictive distributions. The projection of the predictive distribution allows to find the optimal parameter distribution $\pi^{*}$ minimising the KL discrepancy among the induced predictive distribution of the submodel and the reference model one:
\
\begin{equation} 
\pi^{*}\quad \text{s.t.} \quad \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi^{*}}(\tilde{y})]=\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
Consider now the predictive performance of such submodel, a common utility in Bayesian context is the expected log-predictive density defined as:
\
\begin{equation} 
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
\end{equation}
where $p_{t}$ is the distribution of a new observation under the true data generation mechanism. Maximising the elpd of a model is equivalent to minimising the KL-divergence of its predictive distribution from the true data generation mechanism. Therefore the projection corresponds to maximising the elpd using as distribution of the true data generation mechanism the posterior predictive distribution of the reference model, i.e.:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] \leftrightarrow \underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p_{\text{ref}}(\tilde{y}|D)d\tilde{y} 
\end{equation}
Equation \eqref{proj_as_filter} clearly shows the action of the reference model as a filter on our target observations, substituting the unknown true data generation mechanism, that is usually estimated with the data empirical distribution, with the best known approximation that is available, i.e. the reference model posterior predictive distribution.
